\documentclass[10pt]{article}

\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{xcolor}
\usepackage{cancel}
\usepackage{graphicx}
\usepackage{changepage}
\usepackage{circuitikz}
\usepackage{pgfplots}
\usepackage{physics}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage{fontspec}
\usepackage{relsize}
\usepackage{subfig}
\usepackage{todonotes}
\usepackage{multicol, multirow, booktabs}
\usepackage[breakable]{tcolorbox}
\usepackage[inline]{enumitem}

\theoremstyle{definition}
\newtheorem{problem}{Problem}
\newtheorem{soln}{Solution}

\pgfplotsset{compat=newest}
\usetikzlibrary{lindenmayersystems}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning, fit}
\usetikzlibrary{3d, perspective}

\definecolor{incolor}{HTML}{303F9F}
\definecolor{outcolor}{HTML}{D84315}
\definecolor{cellborder}{HTML}{CFCFCF}
\definecolor{cellbackground}{HTML}{F7F7F7}
\newcommand{\ui}{\hat{i}}
\newcommand{\uj}{\hat{j}}
\newcommand{\uk}{\hat{k}}
\newcommand{\ux}{\hat{x}}
\newcommand{\uy}{\hat{y}}
\newcommand{\uz}{\hat{z}}
\newcommand{\primed}[1]{#1^\prime}
\pgfdeclarelayer{background}  
\pgfsetlayers{background,main}
\AtBeginDocument{\RenewCommandCopy\qty\SI}

\makeatletter
\newcommand{\boxspacing}{\kern\kvtcb@left@rule\kern\kvtcb@boxsep}
\makeatother
\newcommand{\prompt}[4]{
    \ttfamily\llap{{\color{#2}[#3]:\hspace{3pt}#4}}\vspace{-\baselineskip}
}

\newcommand{\thevenin}[2]{
  \begin{center}
    \begin{circuitikz} \draw
      (0,0) -- (2,0) to[battery1, l_=$V_{Th}\eq#1$] (2,2) 
      to[resistor, l_=$R_{Th}\eq#2$] (0,2)
      ;
      \draw [o-] (-.07,2.079);
      \draw [o-] (-.07,0.079);
    \end{circuitikz}
  \end{center}
}

\newcommand{\norton}[2]{
  \begin{center}
    \begin{circuitikz} \draw
      (0,0) -- (3,0) to[american current source, l_=$I_{N}\eq#1$] (3,2) -- (0,2) (2,0)
      to[resistor, l=$R_{N}\eq#2$] (2,2)
      ;
      \draw [o-] (-.07,2.079);
      \draw [o-] (-.07,0.079);
    \end{circuitikz}
  \end{center}
}

\newcommand{\highlight}[1]{\colorbox{yellow}{$\displaystyle #1$}}

\newcommand{\ti}[1]{\widetilde{#1}}

\newfontface{\Kaufmann}{Kaufmann}
\DeclareTextFontCommand{\kf}{\Kaufmann}
\newcommand{\scriptr}{\fontsize{12pt}{12pt}\kf{r}}

\newfontface{\KaufmannB}{Kaufmann Bd BT}
\DeclareTextFontCommand{\kfb}{\KaufmannB}
\newcommand{\bscriptr}{\fontsize{12pt}{12pt}\kfb{r}}

\newcommand{\bv}[1]{\mathbf{#1}}

\title{Math 3310H: Assignment I}
\author{Jeremy Favro (0805980) \\ Trent University, Peterborough, ON, Canada}
\date{\today}

\begin{document}
\maketitle

% PROBLEM 1
\begin{problem}
Define a relation $\mathbb{R}\cross\mathbb{R}$ by $(a,b)\sim(c,d)$ if $2(a-c)-3(b-d)=0$
\begin{enumerate}[label=(\alph*)]
  \item Show that $\sim$ is an equivalence relation on $\mathbb{R}$.
  \item Give an example of two pairs $(a,b),(c,d)\in\mathbb{R}\cross\mathbb{R}$, which lie in the same
        equivalence class, and two pairs that don't.
  \item This equivalence relation partitions the 2D plane $\mathbb{R}\cross\mathbb{R}$ into subregions.
        What does the equivalence class $(a,b)$ look like as a region of the plane?
\end{enumerate}
\end{problem}
\begin{soln}
  \begin{enumerate}[label=(\alph*)]
    \item For $\sim$ to be an equivalence relation it must satisfy the following properties for a set $S$ (proofs included)
          \begin{enumerate}[label=(\roman*)]
            \item Reflexivity: $x\sim x\,\forall x\in S$.
                  \begin{proof}
                    Let $(a,b)\in\mathbb{R}\cross\mathbb{R}$, then
                    \begin{align*}
                               & (a,b)\stackrel{?}{\sim}(a,b) \\
                      \implies & 2(a-a)-3(b-b) = 0
                    \end{align*}
                    Which satisfies our relation as defined. Therefore the relation is reflexive. \qedhere
                  \end{proof}
            \item Symmetry: $x\sim y\implies  y\sim x\,\forall x,y\in S$
                  \begin{proof}
                    Let $(a,b), (c,d)\in\mathbb{R}\cross\mathbb{R}$, then
                    \begin{align*}
                               & (a,b)\sim(c,d)          \\
                      \implies & 2(a-c)-3(b-d) = 0       \\
                      \implies & 2(a-c) = 3(b-d)         \\
                      \implies & -2(a-c) = -3(b-d)       \\
                      \implies & 2(c-a) = 3(d-b)         \\
                      \implies & 2(c-a)-3(d-b) =0        \\
                      \implies & (c,d)\sim(a,b) \qedhere
                    \end{align*}
                  \end{proof}
            \item Transitivity: $x\sim y\sim z\implies  x\sim z\,\forall x,y,z\in S$
                  \begin{proof}
                    Let $(a,b), (c,d), (e,f)\in\mathbb{R}\cross\mathbb{R}$, then
                    \begin{align*}
                               & (a,b)\sim(c,d)    \\
                      \implies & 2(a-c)-3(b-d) = 0 \\
                    \end{align*}
                    and
                    \begin{align*}
                               & (c,d)\sim(e,f)    \\
                      \implies & 2(c-e)-3(d-f) = 0 \\
                    \end{align*}
                    so
                    \begin{align*}
                               & 2(a-c)-3(c-d) + 2(c-e)-3(d-f)  = 0   \\
                      \implies & 2(a-c + c-e)-3(b-d + d-f)  = 0 \\
                      \implies & 2(a-e)-3(b-f)  = 0 \\
                      \implies & (a,b)\sim(e,f) \qedhere\\
                    \end{align*}
                  \end{proof}
          \end{enumerate}
          Therefore $\sim$ is an equivalence relation on $\mathbb{R}\cross\mathbb{R}$.
    \item For representative element $(1,1)$ we get that for an element $(a,b)\in \mathbb{R}\cross\mathbb{R}$ 
    to belong to the associated equivalence class we must have
    $$2(1-a)-3(1-b)=0$$
    which can be rearranged to obtain
    $$a=-\frac{1-3b}{2}$$
    so for $b=\pm 1$ we get two members of the equivalence class represented by $(1,1)$ under $\sim$, $(1,1)$ and $(-2,1)$. 
    The elements $(\pi, e)$ and $(\phi, i^i)$ where $\pi,e$ take on their usual definitions, $\phi$ is the golden ratio and $i^i$ is, interestingly,
    both transcendental \emph{and} real!
    \item The equivalence class with representative $(a,b)$ is the set $E=\left\{(x,y)\in \mathbb{R}\cross\mathbb{R} | x\sim (a,b)\right\}$. This gives the equation
    $$2(a-x)-3(b-y)=0\implies y=\frac{2(a-x)-3b}{-3}$$
    so the class looks like a line with slope $2/3$ and y-intercept $b-2a/3$
  \end{enumerate}
\end{soln}

% PROBLEM 2
\begin{problem}
For each of the following sets $S$, determine whether $S$ is closed under
addition modulo $n$, or multiplication modulo $n$, or both or neither.
(Addition and multiplication modulo $n$ are defined in Exercise Set 2).
\begin{enumerate}[label=(\alph*)]
  \item $S=\left\{0,4,8,12\right\},n=16$.
  \item $S=\left\{0,3,6,9,12\right\},n=15$.
  \item $S=\left\{1,2,3,4\right\},n=5$.
  \item $S=\left\{0,2,3,4,6,8,9,10\right\},n=12$.
  \item $S=\left\{1,5,7,11\right\},n=12$.
\end{enumerate}
\end{problem}
\begin{soln}
\end{soln}

% PROBLEM 3
\begin{problem}
Determine whether the given binary operation $*$ is commutative,
associative, both or neither. Justify your answers with proof.
\begin{enumerate}[label=(\alph*)]
  \item The operation $*$ on $\mathbb{Z}$ given by $a*b=a+b+ab$
  \item The operation $*$ on $\mathbb{R}$ given by $a*b=a+b-ab$
  \item The operation $*$ on $\mathbb{R}$ given by $a*b=a+2ab$
  \item The operation $*$ on $\mathbb{Z}\cross\mathbb{Z}$ given by $(a,b)*(c,d)=(ad+bc, bd)$
  \item The operation $*$ on $\mathbb{Z}\cross\mathbb{Z}$ given by $(a,b)*(c,d)=(ad,bc)$
\end{enumerate}
\end{problem}
\begin{soln}
\end{soln}

% PROBLEM 4
\begin{problem}
Let $S$ be a nonempty set. A binary algebraic structure $(S, *)$ is called
a semigroup if $*$ is associative.
\begin{enumerate}[label=(\alph*)]
  \item Let $S$ be the set of positive rational numbers. Show that $(S, *)$
        is a commutative semigroup if
        $$a*b=\frac{ab}{a+b}$$
        (the usual operations on the right) for all $a,b\in S$
  \item Let $S$ be a set containing more than one element. Define
        $$a*b=b$$
        for all $a,b\in S$. Show that $(S,*)$ is a noncommutative semigroup with no identity element.
\end{enumerate}
\end{problem}
\begin{soln}
\end{soln}
\end{document}