\documentclass[10pt]{article}

\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath,amsthm,amssymb,bm}
\usepackage{xcolor}
\usepackage{cancel}
\usepackage{graphicx}
\usepackage{changepage}
\usepackage{circuitikz}
\usepackage{pgfplots}
\usepackage{minted}
\usepackage{physics}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage{fontspec}
\usepackage{relsize}
\usepackage{subfig}
\usepackage{todonotes}
\usepackage{contour}
\usepackage{multicol, multirow, booktabs}
\usepackage[breakable]{tcolorbox}
\usepackage[inline]{enumitem}

\theoremstyle{definition}
\newtheorem{problem}{Problem}
\newtheorem{soln}{Solution}

\pgfplotsset{compat=newest}
\usetikzlibrary{lindenmayersystems}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{positioning, fit}
\usetikzlibrary{3d, perspective}
\usetikzlibrary{math}
\usetikzlibrary{fadings}
\usetikzlibrary{angles,quotes} 

\definecolor{incolor}{HTML}{303F9F}
\definecolor{outcolor}{HTML}{D84315}
\definecolor{cellborder}{HTML}{CFCFCF}
\definecolor{cellbackground}{HTML}{F7F7F7}
\newcommand{\ui}{\hat{i}}
\newcommand{\uj}{\hat{j}}
\newcommand{\uk}{\hat{k}}
\newcommand{\ux}{\hat{\mathbf{x}}}
\newcommand{\uy}{\hat{\mathbf{y}}}
\newcommand{\uz}{\hat{\mathbf{z}}}
\newcommand{\uv}[1]{\hat{\bv{{#1}}}}
\newcommand{\pr}[1]{{#1^\prime}}
\newcommand{\pd}[2][]{{\frac{\partial^{#1}}{\partial {{#2}^{#1}}}}}
\newcommand{\dif}[2][]{{\frac{d^{#1}}{d{{#2}^{#1}}}}}
\newcommand{\grd}{\nabla}

\pgfdeclarelayer{background}  
\pgfsetlayers{background,main}
\AtBeginDocument{\RenewCommandCopy\qty\SI}

\makeatletter
\newcommand{\boxspacing}{\kern\kvtcb@left@rule\kern\kvtcb@boxsep}
\makeatother
\newcommand{\prompt}[4]{
    \ttfamily\llap{{\color{#2}[#3]:\hspace{3pt}#4}}\vspace{-\baselineskip}
}

\newcommand{\thevenin}[2]{
  \begin{center}
    \begin{circuitikz} \draw
      (0,0) -- (2,0) to[battery1, l_=$V_{Th}\eq#1$] (2,2) 
      to[resistor, l_=$R_{Th}\eq#2$] (0,2)
      ;
      \draw [o-] (-.07,2.079);
      \draw [o-] (-.07,0.079);
    \end{circuitikz}
  \end{center}
}

\newcommand{\norton}[2]{
  \begin{center}
    \begin{circuitikz} \draw
      (0,0) -- (3,0) to[american current source, l_=$I_{N}\eq#1$] (3,2) -- (0,2) (2,0)
      to[resistor, l=$R_{N}\eq#2$] (2,2)
      ;
      \draw [o-] (-.07,2.079);
      \draw [o-] (-.07,0.079);
    \end{circuitikz}
  \end{center}
}

\DeclareMathOperator{\Div}{div}
\DeclareMathOperator{\Curl}{curl}
\DeclareMathOperator{\sgn}{sgn}

\newcommand{\highlight}[1]{\colorbox{yellow}{$\displaystyle #1$}}
\newcommand{\ti}[1]{\widetilde{#1}}
\renewcommand{\div}[1]{\bv{\nabla}\cdot #1}
\renewcommand{\curl}[1]{\bv{\nabla}\times #1}

\newfontface{\Kaufmann}{Kaufmann}
\DeclareTextFontCommand{\kf}{\Kaufmann}
\newcommand{\scriptr}{\fontsize{12pt}{12pt}\kf{r}\fontsize{10pt}{10pt}}

\newfontface{\KaufmannB}{Kaufmann Bold}
\DeclareTextFontCommand{\kfb}{\KaufmannB}
\newcommand{\bscriptr}{\fontsize{12pt}{12pt}\kfb{r}\fontsize{10pt}{10pt}}
\newcommand{\justif}[2]{&{#1}&\text{#2}}
\newcommand{\bv}[1]{\bm{\mathrm{#1}}}

\title{Physics 3200Y: Assignment IX}
\author{Jeremy Favro (0805980) \\ Trent University, Peterborough, ON, Canada}
\date{\today}

\begin{document}
\maketitle

% PROBLEM 1
\begin{problem}
A metal bar of mass $m$ slides frictionlessly on two parallel conducting rails a distance $\ell$ apart.
A resistor $R$ is connected across the rails and a uniform magnetic field $\bv{B}$, pointing into the page,
fills the entire region.
\begin{enumerate}[label=(\alph*)]
  \item If the bar moves to the right with speed $v$, what is the current in the resistor? In what direction does it flow?
  \item What is the magnetic force on the bar? In what direction?
  \item If the bar starts out with speed $v_0$ at time $t=0$, and is left to slide, what is its speed at a later time $t$?
  \item The initial kinetic energy of the bar was, of course, $\frac{1}{2}mv_0^2$. Check that the energy delivered to the
        resistor is exactly $\frac{1}{2}mv_0^2$.
\end{enumerate}
\end{problem}
\begin{soln}~
  \begin{enumerate}[label=(\alph*)]
    \item We know that the current in a purely resistive circuit is related to the flux, and thereby the EMF, by
          $$-\frac{d\phi}{dt}=\epsilon=IR.$$
          Since the flux here is the field times the area of the loop, say the bar is at a position $x(t)$. Then
          $$\phi=B\ell x\implies IR=-B\ell\frac{dx}{dt}\implies I=-\frac{B\ell v}{R}.$$
          Where the negative sign indicates direction. Since the field is pointing into the page we can see that, using the
          RHR, the current is traveling downward through the resistor.
    \item Magnetic force is given by $\bv{f}=dq \bv{v}_{charges}\times \bv{B}$. Here since $\bv{v}_{charges}$ and $\bv{B}$ are at right
          angles to one another the cross product resolves to a regular product of magnitudes which we can integrate along the bar
          to find the force on the whole thing,
          $$\bv{F}=\int_{0}^{\ell}dq v_{charges} B(-\ux)=\int_{0}^{\ell}\frac{dq}{dt}Bd\ell'(-\ux)=I\ell B(-\ux)=-\frac{B^2\ell^2 v}{R}\ux.$$
          where we haven't picked up the sign on the current because we're considering magnitudes. This says that the force is leftward.
    \item The equation of motion here is
          $$m\ddot{x}=m\frac{dv}{dt}=-\frac{B^2\ell^2 v}{R}\implies v(t)=v_0\exp\left(-\frac{B^2\ell^2}{mR}t\right)$$
    \item The energy dissipated by the resistor is the power, given by $I^2R$:
          $$\frac{dW}{dt}=I^2 R=\frac{B^2\ell^2 v^2}{R}=\frac{B^2\ell^2}{R}v_0^2\exp\left(-2\frac{B^2\ell^2}{mR}t\right).$$
          Integrating this from $t=0$ to steady-state then
          \begin{align*}
            W & =\frac{B^2\ell^2}{R}v_0^2\int_0^\infty \exp\left(-2\frac{B^2\ell^2}{mR}t\right)dt                             \\
              & =-\frac{B^2\ell^2}{R}v_0^2\frac{mR}{2B^2\ell^2}\left[\exp\left(-2\frac{B^2\ell^2}{mR}t\right)\right]_0^\infty \\
              & =-\frac{1}{2}mv_0^2\left[e^{-\infty}-e^{0}\right]=\frac{1}{2}mv_0^2.
          \end{align*}
  \end{enumerate}
\end{soln}

% PROBLEM 2
\begin{problem}
An alternating current $I=I_0\cos\omega t$ flows down a long straight wire, and returns along a coaxial conducting tube
of radius $a$.
\begin{enumerate}[label=(\alph*)]
  \item In what \emph{direction} does the induced electric field point (radial, circumferential, longitudinal)?
  \item Assuming that the field goes to zero as $s \to \infty$, find $\bv{E}(s,t)$.
\end{enumerate}
\end{problem}
\begin{soln}
  \begin{enumerate}[label=(\alph*)]
    \item If we consider an Amp\`erian loop enclosing only the central conductor we see that the magnetic field
          is entirely in the $\uv{\phi}$ direction in cylindrical coordinates. Then since
          $$\curl{\bv{E}}=-\frac{d\bv{B}}{dt}$$
          $\curl{\bv{E}}$ must also be in $\uv{\phi}$. Looking at the $\uv{\phi}$ term for curl in
          cylindrical we see that
          $$\curl{\bv{v}}=\dots+\left[\frac{\partial v_s}{\partial z}-\frac{\partial v_z}{\partial s}\right]\uv{\phi}+\dots.$$
          Since the wire is long there shouldn't be any $z$ dependence so $\bv{E}$ should only have a $\uz$ component which varies with
          $s$.
          \todo{Contour diagrams\dots}
    \item Taking a new contour of length $\ell$ which slices the outer conductor we note that using our definition of flux,
          $$\int (\curl{\bv{E}})\cdot d\bv{a}=\oint \bv{E}\cdot d\ell'\uz=-\frac{d}{dt}\int \bv{B}\cdot d\bv{a}$$
          where the first equality comes from Stoke's (curl) theorem. We can also use Ampere's law to write
          $$\oint \bv{B}\cdot d\bv{\ell}'=2\pi s' B=\mu_0 I_0\cos\omega t.$$
          Plugging this into the chain of equalities for the electric field gives that
          $$E\ell=-\frac{d}{dt}\int_0^\ell\int_s^a \frac{\mu_0 I_0}{2\pi s'}\cos\omega t ds'dz'$$
          and so
          $$E=\frac{\mu_0 I_0 \omega}{2\pi}\ln\left(\frac{a}{s}\right)\sin\omega t.$$
  \end{enumerate}
\end{soln}

% PROBLEM 3
\begin{problem}
An infinite cylinder of radius $R$ carries a uniform surface charge $\sigma$.
We propose to set it spinning about its axis, at a final angular velocity $\omega_f$. How much work will this take per unit length?
Do it two ways, and compare your answers:
\begin{enumerate}[label=(\alph*)]
  \item Find the magnetic field and the induced electric field (in the quasistatic approximation),
        inside and outside the cylinder, in terms of $\omega$, $\dot{\omega}$, and $s$. Calculate the torque you must
        exert, and from that obtain the work done per unit length ($W=\int Nd\phi$).
  \item Use Eq. 7.44 ($W=\frac{1}{2\mu_0}\int_{\mathbb{R}}B^2d\tau$) to determine the energy stored in the resulting magnetic field.
\end{enumerate}
\end{problem}
\begin{soln}~
  \begin{enumerate}[label=(\alph*)]
    \item When we set this system rotating it becomes a solenoid. We know the field of a solenoid to be
          $$\bv{B}=\begin{cases}
              \mu_0\sigma\omega R\uz & s<R  \\
              0                      & s> R
            \end{cases}$$
          where $\sigma\omega R$ is the surface current density. Since, for a contour with area in $\uz$
          $$\oint \bv{E}\cdot d\bv{\ell}'=-\frac{d}{dt}\int\bv{B}\cdot d\bv{a}$$
          and we expect that the electric field is entirely in $\uv{\phi}$, as is $d\bv{\ell}'$,
          $$E2\pi s=-\frac{d}{dt}\int_{0}^{s}s'ds'\int_{0}^{2\pi}d\phi' \mu_0\sigma\omega R=-2\pi\mu_0\sigma\dot{\omega}\frac{R s^2}{2}$$
          so
          $$\bv{E}=-\mu_0\sigma\dot{\omega}\frac{R s}{2}\uv{\phi}$$
          inside. To do outside we change the limits of the integrals,
          $$E2\pi s=-\frac{d}{dt}\int_{0}^{R}s'ds'\int_{0}^{2\pi}d\phi' \mu_0\sigma\omega R=-2\pi\mu_0\sigma\dot{\omega}\frac{R^3}{2}$$
          so
          $$\bv{E}=-\mu_0\sigma\dot{\omega}\frac{R^3}{2s}$$
          outside. At the surface then (I think) the electric field is
          $$\bv{E}=-\mu_0\sigma\dot{\omega}\frac{R^2}{2}\uv{\phi}.$$
          Torque then is
          $$\bv{\tau}=\bv{r}\times\bv{F}=R(2\pi R\ell \sigma)(-\mu_0\sigma\dot{\omega}\frac{R^2}{2})\uz$$
          since the total charge on a segment of length $\ell$ is $2\pi R\ell \sigma$.
          This works out to
          $$\bv{\tau}=-\pi\ell\mu_0\sigma^2\dot{\omega}R^4\uz.$$
          Work per $\ell$ then is
          $$\frac{W}{\ell}=-\pi\mu_0\sigma^2R^4\int\dot{\omega}d\phi$$
          and
          $$\dot{\omega}d\phi=\frac{d\omega}{dt}\omega dt=\omega d\omega$$
          so
          $$\frac{W}{\ell}=-\pi\mu_0\sigma^2R^4\int_0^{\omega_f}\omega d\omega=-\frac{\pi\mu_0}{2}\sigma^2R^4\omega_f^2.$$
    \item $$W=\frac{1}{2\mu_0}\int_0^{2\pi}\int_{0}^{R}\int_{0}^{\ell}\mu_0^2\sigma^2\omega^2 R^3dz ds d\phi=\pi\mu_0\sigma^2\omega^2 R^4\ell$$
          which gives the same result per unit length as previous except with the sign reversed. The sign change is because this is the work done to bring the field
          up to strength by the person spinning it up and the other (negative) work was the work done by the electric field.
  \end{enumerate}
\end{soln}

% PROBLEM 4
\begin{problem}
Consider a capacitor made of two circular disks with wires connected to the centre of each
disk. The disks have radius $a$ and are separated by $w \ll a$. For concreteness, put one disk
at $z = -w/2$ and the other at $z=w/2$. A current $I(t)$ flows through the capacitor. Assume
that the wires are infinitely thin.
\begin{enumerate}[label=(\alph*)]
  \item In 2D, the continuity equation is $\div{\bv{K}}=-\partial\sigma/\partial t$, with $\sigma$ the surface charge density
        and $\bv{K}$ the surface current density. Assume that charge flows onto the disks in such a
        way that $\sigma(t)$ is independent of position on each disk. To understand what this means,
        think about filling up a basin from a tap. If the liquid coming out of the tap is viscous
        (like honey), the height of the liquid in the basin will be higher near the tap than far
        away from it. If the liquid is not viscous (like water), the height of the surface will
        remain the same everywhere in the basin. Here, we assume that electrons spread out
        in such a way as to keep $\sigma$ uniform at all times. What must $\bv{K}$ be for this to happen?
        Note that you have two boundary conditions:
        \begin{enumerate}[label=(\roman*)]
          \item The total current flowing onto the disk is $I$;
          \item The current density at the edge of the disk (at $s = a$; $a$ is the radius of the disk)
                is zero, since no current can flow through the edge.
        \end{enumerate}
        Your solution for $\bv{K}$ will diverge at $s = 0$, but this is a consequence of assuming the
        wires carrying the current onto the disks are infinitely thin.
  \item What is the magnetic field produced by $\bv{K}$? Set up the integral for the field at a general
        position $\bv{r}$, but then evaluate it for the simple case $\bv{r}=(0,0,z)$ with $-w/2<z<w/2$.
  \item What is the magnetic field produced by the displacement current? You should be able
        to obtain this for any point that's not too close to the edge of the disk.
\end{enumerate}
\end{problem}
\begin{soln}~
  \begin{enumerate}[label=(\alph*)]
    \item Here we expand the continuity equation using the definition of the divergence in cylindrical coordinates.
          $$\div{\bv{K}}=\frac{1}{s}\frac{\partial (K_ss)}{\partial s}=-\frac{\partial \sigma}{\partial t}$$
          and so
          $$K_s=-\frac{\partial \sigma}{\partial t}\frac{s}{2}+\frac{C(t)}{s}$$
          and applying the second boundary condition,
          $$K_s(a)=0=-\frac{\partial \sigma}{\partial t}\frac{a}{2}+\frac{C(t)}{a}.$$
          And so
          $$K_s(s)=-\frac{\partial \sigma}{\partial t}\frac{s}{2}+\frac{\partial \sigma}{\partial t}\frac{a^2}{2s}=\frac{1}{2}\frac{\partial \sigma}{\partial t}\left[\frac{a^2}{s}-s\right].$$
          The only reason to apply the first boundary conditon is to determine $\sigma$ in terms of $I$. To do this note that
          $$\frac{\partial \sigma}{\partial t}=\frac{1}{\pi a^2}\frac{\partial Q}{\partial t}=\frac{1}{\pi a^2}I$$
          so
          $$\bv{K}=\frac{I}{2\pi}\left[\frac{1}{s}-\frac{s}{a^2}\right]\uv{s}$$
    \item The general form of the magnetic field due to a surface current is
          $$\bv{B}(\bv{r})=\frac{\mu_0}{4\pi}\int \frac{\bv{K}(\bv{r}')\times \uv{\bscriptr}}{\scriptr^2}.$$
          Here
          $\bv{r}=(x,y,z)$, $\bv{r}'=(x',y',\pm w/2)$, so $\bscriptr=(x-x',y-y', z\mp w/2)$. Since
          $$\bv{K}(s')=\frac{1}{2}\frac{\partial \sigma}{\partial t}\left[\frac{a^2}{s'}-s'\right]\uv{s}=\frac{1}{2}\frac{\partial \sigma}{\partial t}\left[\frac{a^2}{s'}-s'\right]\left(\cos\phi\ux+\sin\phi\uy\right)$$
          the cross product comes out to
          $$
            \bv{K}\times \bv{\scriptr}
            =\begin{vmatrix}
              \ux        & \uy        & \uz      \\
              K\cos\phi' & K\sin\phi' & 0        \\
              x-x'       & y-y'       & z\mp w/2
            \end{vmatrix}=
            \left[K\left(z\mp w/2\right)\sin\phi'\right]\ux
            -\left[K\left(z\mp w/2\right)\cos\phi'\right]\uy
            \left[K(y-y')\cos\phi'-K(x-x')\sin\phi'\right]\uz.
          $$
          When we take $\bv{r}=(0,0,z)$ and note that $x'=s'\cos\phi'$ and $y'=s'\sin\phi'$ this becomes
          $$
            \bv{K}\times \bv{\scriptr}
            =
            \left[K\left(z\mp w/2\right)\sin\phi'\right]\ux
            -\left[K\left(z\mp w/2\right)\cos\phi'\right]\uy
            \left[-Ks'\sin\phi'\cos\phi'+Ks'\cos\phi'\sin\phi'\right]\uz.
          $$
          Obviously the $\uz$ term is zero. When we integrate the other two over the entire disc(s) the $\sin\phi'$ and
          $\cos\phi'$ will go to zero. This means, on axis, $\bv{B}(\bv{r})=\bv{0}$.
    \item Here
          $$\curl{\bv{B}}=\mu_0\bv{J}+\mu_0\epsilon_0\frac{\partial \bv{E}}{\partial t}.$$
          Since we are considering the region between the plates, which is vaccuum, $\bv{J}=0$ and so
          $$\curl{\bv{B}}=\mu_0\epsilon_0\frac{\partial \bv{E}}{\partial t}=\frac{\mu_0 I}{\pi a^2}\uz$$
          where we've recalled that $\bv{E}=\sigma/\epsilon_0$ for a parallel plate capacitor and used the current definition we worked out in (a).
          Since we expect the RHS to be in $\uz$ only so should the LHS meaning
          $$\curl{\bv{B}}=\frac{1}{s}\left[\frac{\partial (s B_\phi)}{\partial s}-\frac{\partial B_s}{\partial \phi}\right]\uz=\frac{\mu_0 I}{\pi a^2}\uz$$
          and since the plates are symmetric in $\phi$ the $\phi$ derivative should be zero so
          $$\bv{B}=\frac{\mu_0 Is}{2\pi a^2}\uv{\phi}.$$
  \end{enumerate}
\end{soln}
\end{document}